---
title: "Coursera Developing Data Products Final Project"
author: "Gudur Guy"
date: "6/30/2023"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

## Introduction

 - This presentation is one half of the assignemnt of week 4, ***Developing Data Products course from Coursera*** (<https://www.coursera.org/learn/data-products>). 
 
 - The presentation is meant to show information about the second half of the above mentioned assignment (a development project). 

 - The presentation was generated using ***RStudio***(<https://www.rstudio.com>) and ***Slidify***(<http://slidify.org>) framework.

## Application

 - The second half of the mentioned assignment was to develop a web application. The application was named ***City and Highway Mileage Predictor***. 
 
 - An instance is up & running at <https://gudurguy.shinyapps.io/developing-data-products-app/>.
 

 - This application shows an example of making a web application using R and [Shiny](https://shiny.rstudio.com) framework together.
 
 - In the application, two random forest prediction models are generated, based on a car parameters such as manufatures, model, type of drive, number of cylinders and class of the vehicle etc.. The user can play freely with the UI values in order to simulate the parameters of a hypothetical car and be able to predict its Mileage for both City and Highway runs.
 
 - The application and the current presentation source codes can be found at <https://github.com/gudurguy/coursera-developing-data-product-project>. The application comprises 3 files: *ui.R* (UI), *server.R* (backend) and *rfModel.R* (Random Forest predictors).
  
## Dataset

Dataset comes from ggplot2 and is called mpg. The data was extracted from the 1998 to 2008 for 38 popular models of cars.

The description of dataset: 
```{r cars, echo = TRUE}
str(mpg)
```

## Prediction

Two Random Forest prediction models are generated and trained using the *'mpg'* dataset. Two Model because one for City Mileage Prediction and one for Highway Mileage Prediction. The goal of these models is to predict the fuel consumption (***mpg*** variable) based on the rest of the variables:
```{r echo = FALSE}
suppressWarnings(suppressMessages(library(caret)))
suppressWarnings(suppressMessages(library(randomForest)))
set.seed(210915)
```

```{r echo = TRUE}
# Train Random Forest model for predicting "cty" mileage given all other variables in the dataset. 
rfCtyModelFit <- function() {  
  return(
    train(cty ~ ., data = cityTrainingData, method = "rf", trControl = trainControl(method = "cv", 5), ntree = 250)
  )
}

# Train Random Forest model for predicting "hwy" mileage given all other variables in the dataset. 
rfHwyModelFit <- function() {
  return(
    train(hwy ~ ., data = hwyTrainingData, method = "rf", trControl = trainControl(method = "cv", 5), ntree = 250)  
  )
}

# "cty" mileage prediction function given user selected variables
rfCtyPrediction <- function(modelFit, predictors) {
  return(
    predict( modelFit, newdata = predictors)  
  )
}

# "hwy" mileage prediction function given user selected variables
rfHwyPrediction <- function(modelFit, predictors) {
  return(
    predict( modelFit, newdata = predictors)  
  )
}
```